\chapter{Set Interface}
\label{ch:set-interface}

\begin{cluster}
\label{grp:prmbl:set-interface::interface}

\begin{preamble}
\label{prmbl:set-interface::interface}
The \sml{SET} interface specifies an unordered collection of items. Sets do not
contain duplicates, and are not polymorphic: the type of their elements is
given by the \sml{Key} substructure.

\end{preamble}
\end{cluster}


\section{Summary}
\label{sec:set-interface::summary}

\begin{cluster}
\label{grp:grm:set-interface::signature}

\begin{gram}
\label{grm:set-interface::signature}
\begin{verbatim}
signature SET =
sig
  structure Key : EQKEY
  structure Seq : SEQUENCE

  type t
  type set = t

  val size : set -> int
  val toString : set -> string
  val toSeq : set -> Key.t Seq.t

  val empty : unit -> set
  val singleton : Key.t -> set
  val fromSeq : Key.t Seq.t -> set

  val find : set -> Key.t -> bool
  val insert : set * Key.t -> set
  val delete : set * Key.t -> set

  val filterKey : (Key.t -> bool) -> set -> set

  val reduceKey : (Key.t * Key.t -> Key.t) -> Key.t -> set -> Key.t
  val iterateKey : ('a * Key.t -> 'a) -> 'a -> set -> 'a

  val union : set * set -> set
  val intersection : set * set -> set
  val difference : set * set -> set

  val $ : Key.t -> set
end
\end{verbatim}

\end{gram}
\end{cluster}


\section{Substructures}
\label{sec:set-interface::substructures}

\begin{cluster}
\label{grp:grm:set-interface::structure}

\begin{gram}
\label{grm:set-interface::structure}
\begin{verbatim}
structure Key : EQKEY
\end{verbatim}
The \sml{Key} substructure defines the type of elements in a set, which may be
compared for equality.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::sequence}

\begin{gram}
\label{grm:set-interface::sequence}
\begin{verbatim}
structure Seq : SEQUENCE
\end{verbatim}
The \sml{Seq} substructure defines the underlying sequence type, so that we
may convert sets to and from sequences.

\end{gram}
\end{cluster}


\section{Types}
\label{sec:set-interface::types}

\begin{cluster}
\label{grp:grm:set-interface::type}

\begin{gram}
\label{grm:set-interface::type}
\begin{verbatim}
type t
type set = t
\end{verbatim}
The abstract set type. The alias \sml{set} is for readability in the
signature.

\end{gram}
\end{cluster}


\section{Values}
\label{sec:set-interface::values}

\begin{cluster}
\label{grp:grm:set-interface::size}

\begin{gram}[size]
\label{grm:set-interface::size}
\begin{verbatim}
val size : set -> int
\end{verbatim}
The number of elements in a set.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::tostring}

\begin{gram}[toString]
\label{grm:set-interface::tostring}
\begin{verbatim}
val toString : set -> string
\end{verbatim}
Evaluates to a string representation of the set. Each element is converted to
a string via \sml{Key.toString}.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::toseq}

\begin{gram}[toSeq]
\label{grm:set-interface::toseq}
\begin{verbatim}
val toSeq : set -> Key.t Seq.t
\end{verbatim}
Return a sequence of all keys in a set. The ordering of the elements in the
returned sequence is implementation-defined.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::empty}

\begin{gram}[empty]
\label{grm:set-interface::empty}
\begin{verbatim}
val empty : unit -> set
\end{verbatim}
Construct the empty set.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:gr:set-interface:singleton}

\begin{gram}[singleton]
\label{gr:set-interface:singleton}
\begin{verbatim}
val singleton : Key.t -> set
\end{verbatim}
Construct the singleton set containing only the provided key.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::fromseq}

\begin{gram}[fromSeq]
\label{grm:set-interface::fromseq}
\begin{verbatim}
val fromSeq : Key.t Seq.t -> set
\end{verbatim}
Return the set of all elements of a sequence.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::find}

\begin{gram}[find]
\label{grm:set-interface::find}
\begin{verbatim}
val find : set -> Key.t -> bool
\end{verbatim}
\sml{find s k} returns whether or not $k$ is a member of the set $s$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::insert}

\begin{gram}[insert]
\label{grm:set-interface::insert}
\begin{verbatim}
val insert : set * Key.t -> set
\end{verbatim}
\sml{insert (s, k)} evaluates to the set $s \cup \{k\}$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::delete}

\begin{gram}[delete]
\label{grm:set-interface::delete}
\begin{verbatim}
val delete : set * Key.t -> set
\end{verbatim}
\sml{delete (s, k)} evaluates to the set $s \setminus \{k\}$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::filterkey}

\begin{gram}[filterKey]
\label{grm:set-interface::filterkey}
\begin{verbatim}
val filterKey : (Key.t -> bool) -> set -> set
\end{verbatim}
\sml{filterKey p s} evaluates to the set $\{x \in s \mathbin| p(x) \}$

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::reducekey}

\begin{gram}[reduceKey]
\label{grm:set-interface::reducekey}
\begin{verbatim}
val reduceKey : (Key.t * Key.t -> Key.t) -> Key.t -> set -> Key.t
\end{verbatim}
\sml{reduceKey f b s} is logically equivalent to \sml{Seq.reduce f b (toSeq s)}.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::iteratekey}

\begin{gram}[iterateKey]
\label{grm:set-interface::iteratekey}
\begin{verbatim}
val iterateKey : ('a * Key.t -> 'a) -> 'a -> set -> 'a
\end{verbatim}
\sml{iterateKey f b s} is logically equivalent to \sml{Seq.iterate f b (toSeq s)}.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::union}

\begin{gram}[union]
\label{grm:set-interface::union}
\begin{verbatim}
val union : set * set -> set
\end{verbatim}
\sml{union a b} evaluates to the set $a \cup b$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::intersection}

\begin{gram}[intersection]
\label{grm:set-interface::intersection}
\begin{verbatim}
val intersection : set * set -> set
\end{verbatim}
\sml{intersection a b} evaluates to the set $a \cap b$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::difference}

\begin{gram}[difference]
\label{grm:set-interface::difference}
\begin{verbatim}
val difference : set * set -> set
\end{verbatim}
\sml{difference a b} evaluates to the set $a \setminus b$.

\end{gram}
\end{cluster}

\begin{cluster}
\label{grp:grm:set-interface::alias}

\begin{gram}[\$]
\label{grm:set-interface::alias}
\begin{verbatim}
val $ : Key.t -> set
\end{verbatim}
An alias for \ref{gr:set-interface:singleton}.

\end{gram}
\end{cluster}

